---
title: "benchmarking"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{benchmarking}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE
)
```

## Overview

This Vignette compares variouse I/O and serialization approaches in R.

The packages compared are:

+ Base read.csv/write.csv
+ Base readRDS/saveRDS
+ fread/fwrite from the [data.table](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html) package
+ write_feather / read_feather from the [feather](https://cran.r-project.org/web/packages/feather/index.html) package
+ qsave / qwrite from [qs](https://github.com/traversc/qs) package
+ read.fst / write.fst from [fst](https://github.com/fstpackage/fst) package

## Preliminary Notes:

+ **fst** only works with data.frames / tabular data but it does conserve classes (including dates)
+ Data frames stored in the **fst** format have full random access, both in column and rows

## My Choice - qs

Inspired by the fst package, qs uses a similar block-compression design using 
either the lz4 or zstd compression libraries. It differs in that it applies a 
more general approach for attributes and object references.

saveRDS and readRDS are the standard for serialization of R data, but these 
functions are not optimized for speed. On the other hand, fst is extremely fast,
but only works on data.frame’s and certain column types.

qs is both extremely fast and general: it can serialize any R object like saveRDS
and is just as fast and sometimes faster than fst.

## Benchmarking / Profiling


```{r setup, echo=FALSE}
# library(easyr)
library(qs)
library(fst)
library(data.table)
library(readr)
library(feather)
library(bench)
library(knitr)
library(rmarkdown)
library(kableExtra)
library(dplyr)
library(purrr)
library(forcats)
library(ggplot2)
library(microbenchmark) # remotes::install_github("joshuaulrich/microbenchmark")
```

```{r sample_data}
dat <- data.frame(x = rnorm(5e6), 
                  y = sample(5e6), 
                  z = sample(letters,5e6, replace = TRUE),
                  l = c(rep(TRUE, 5e6/2), rep(FALSE, 5e6/2)))

dir <- tempdir()

base_file <- fs::path(dir, "dat_base.csv")
rds_file <- fs::path(dir, "dat_rds.RDS")
fwrite_file <- fs::path(dir, "dat_fwrite.csv")
feather_file <- fs::path(dir, "dat_feather.feather")
qs_file <- fs::path(dir, "dat_qs.qs")
fst_file <- fs::path(dir, "dat_fst.fst")

save_bench <- bench::mark(
  base = write.csv(dat, base_file),
  rds = saveRDS(dat, rds_file),
  fwrite = data.table::fwrite(dat, fwrite_file),
  feather = write_feather(dat, feather_file),
  qs = qs::qsave(dat, qs_file, preset = "fast", algorithm = "lz4", 
                 compress_level = 100, shuffle_control = 0, 
                 nthreads = 8, check_hash = FALSE),
  fst = fst::write.fst(dat, fst_file), 
  check = FALSE,
  iterations = 1
)

read_bench <- bench::mark(
  base = read.csv(base_file),
  rds = readRDS(rds_file),
  fread = data.table::fread(fwrite_file),
  feather = read_feather(feather_file),
  qs = qs::qread(qs_file, nthreads = 8),
  fst = fst::read.fst(fst_file),
  check = FALSE,
  iterations = 1
)
```

```{r}
summary(save_bench) %>%
  mutate(expression = forcats::fct_reorder(as.character(expression), min, .desc = TRUE)) %>%
  as_bench_mark() %>%
  select_if(purrr::negate(is.list)) %>%
  select(-`gc/sec`, -n_itr, -n_gc) %>%
  kable(booktabs = TRUE)
```

```{r}
autoplot(save_bench, "boxplot")
```

```{r}
summary(read_bench) %>%
  mutate(expression = forcats::fct_reorder(as.character(expression), min, .desc = TRUE)) %>%
  as_bench_mark() %>%
  select_if(purrr::negate(is.list)) %>%
  select(-`gc/sec`, -n_itr, -n_gc) %>%
  kable(booktabs = TRUE)
```

```{r}
autoplot(read_bench)
```




## Microbenchmark

```{r}
save_bench <- microbenchmark(
  base = write.csv(dat, base_file),
  rds = saveRDS(dat, rds_file),
  fwrite = data.table::fwrite(dat, fwrite_file),
  feather = write_feather(dat, feather_file),
  qs = qs::qsave(dat, qs_file, preset = "fast", algorithm = "lz4", 
                 compress_level = 100, shuffle_control = 0, 
                 nthreads = 8, check_hash = FALSE),
  fst = fst::write.fst(dat, fst_file), 
  times = 1
)

read_bench <- microbenchmark(
  base = read.csv(base_file),
  rds = readRDS(rds_file),
  fread = data.table::fread(fwrite_file),
  feather = read_feather(feather_file),
  qs = qs::qread(qs_file, nthreads = 8),
  fst = fst::read.fst(fst_file),
  times = 1
)

```


```{r}
save_bench
```

```{r}
boxplot(save_bench)
```


```{r}
read_bench
```

```{r}
boxplot(read_bench)
```


```{r}
# fs::dir_delete(dir)
c(
  base_file,
  rds_file,
  fwrite_file,
  feather_file,
  qs_file,
  fst_file
) %>%
  purrr::walk(fs::file_delete)
```


## Example Usage

Here's some example "caching" utility functions utilizing the **qs** package:

```{r eval=FALSE}
#' Write Cache
#'
#' Quick caching utility write function.
#'
#' @param x object to cache
#' @param name name to store object with
#' @param cache_dir path to cache directory
#' @param overwrite logical (default = TRUE)
#'
#' @return x
#' @export
#'
#' @importFrom fs dir_exists dir_create path
#' @importFrom qs qsave
#'
#' @examples
#' mydata <- mtcars
#' write_cache(mydata) # will save to 'cache/mydata'.
#' write_cache(mydata, "mydata-v2", cache_dir = "data/temp") # will save to 'data/temp/mydata-v2'
write_cache <- function(x,
                        name = NULL,
                        cache_dir = "cache",
                        overwrite = TRUE) {

  if (!fs::dir_exists(cache_dir)) fs::dir_create(cache_dir)
  
  if (is.null(name)) name <- deparse(substitute(x))
  
  qs_file <- fs::path(cache_dir, name)
  
  qs::qsave(x, qs_file)
  
  message("Object ", name, "has been cached at '", qs_file, "'.")
  
  return(x)

}

#' Read Cache
#'
#' Quick caching utility read function.
#'
#' @param name name of object to read in.
#' @param cache_dir path to cache directory.
#'
#' @return invisibly attaches object to parent global environment
#' @export
#' @importFrom fs path file_exists
#' @importFrom fst read_fst
read_cache <- function(name = NULL,
                       cache_dir = "cache") {

  if (is.null(name)) name <- deparse(substitute(x))
  
  qs_file <- fs::path(cache_dir, name)
  
  if (!fs::file_exists(qs_file)) {
    stop("File not found in ", cache_dir)
  }
  
  if (fs::file_exists(qs_file)) out <- qs::qread(qs_file)
  
  assign(name, out, envir = .GlobalEnv)
  
  return(invisible())

}

```

An alternative could be to use **fst** and **RDS** depending on what you are trying to save:

```{r eval=FALSE}
write_cache <- function(x,
                        name = NULL,
                        cache_dir = "cache",
                        overwrite = TRUE) {

  if (!dir.exists(cache_dir)) dir.create(cache_dir)

  if (is.null(name)) name <- deparse(substitute(x))

  fst_file <- paste0(file.path(cache_dir, name), ".fst")
  rds_file <- paste0(file.path(cache_dir, name), ".RDS")

  if (tibble::is_tibble(x) || is.data.frame(x)) {

    tryCatch(
      fst::write_fst(x, fst_file),
      error = function(e, ...) saveRDS(x, rds_file)
    )

  } else {

    saveRDS(x, rds_file)

  }

  if (file.exists(fst_file) && file.exists(rds_file)) fs::file_delete(fst_file)

}

read_cache <- function(name,
                       cache_dir = "cache") {

  fst_file <- paste0(file.path(cache_dir, name), ".fst")
  rds_file <- paste0(file.path(cache_dir, name), ".RDS")

  if (!file.exists(fst_file) && !file.exists(rds_file)) stop("File not found in ", cache_dir)

  if (file.exists(fst_file)) out <- read_fst(fst_file)

  if (file.exists(rds_file)) out <- readRDS(rds_file)

  assign(name, out, envir = .GlobalEnv)

  return(invisible())

}
```

